Online Laterality Battery Pre-Processing - short version

### 2020-04-23 battery-preprocessing_v5_DB.Rmd
Done a bit more with outliers on dichotic, but still grappling with RDT data - why some have NA for RT and what to do with them. Need to look at when fresher.

---
title: "Online Laterality Battery Pre-Processing - short version"
author: "Adam Parker + Dorothy Bishop"
date: "20th March 2020"
output:
  pdf_document: default
  html_document: default
---

@@

This version updated on 1st Feb 2020 to make script more reproducible and compact by using functions for repeated operations.

The script pulls in csv files and processes them so that pre-registered analyses can be conducted (pre-registered on the OSF: ). First, questionnaires are analyses followed by visual half field tasks. This is done for time 1 and time 2. A "full" laterality index output is then appended to a file with the demographics. This is then used in subsequent R markdown scripts.

This script analyses the following:

1) Questionnaires
  - Demographics
  - Miles test
  - Porta test
  - Edinburgh Handedness Inventory
  - lexTALE
2) Chimeric face task
3) Rhyme detection task
4) Key Pressing Task (finger dexterity)
5) Summary of demographics for the final sample

@@

- Demographics
- Miles test
- Porta test
- Edinburgh Handedness Inventory
- lexTALE

@@

#NB for present setwd("~/Dropbox/Online LDT/300 Ps")
# libraries 
library(readr)
library(yarrr)
library(dplyr)
library(tidyr)
library(car)
library(ggplot2)
library(knitr)
require(readr)  # for read_csv()
require(dplyr)  # for mutate()
require(tidyr)  # for unnest()
require(purrr)  # for map(), reduce()
library(GGally)
require(stringr) # string matching
library(ggpirate)
library(corrr) #easier correlations!
```

@@

```{r demographics, warning=FALSE}
demo_dat_R <- read_csv("./Demographics_R.csv", na = c("empty", "NA")) # read in data
demo_dat_L <- read_csv("./Demographics_L.csv", na = c("empty", "NA")) # read in data
demo_dat <- rbind(demo_dat_R, demo_dat_L)
# relabel 
demo_dat$subject <- demo_dat$`Participant Public ID`
demo_dat$subject <- as.factor(demo_dat$subject)
# count and print subjects
nsub <-length(levels(demo_dat$subject))
print(paste0("subjects: ", nsub))

# create data frame
demographics <- data.frame(matrix(ncol = 7, nrow = nsub))
  colnames(demographics) <- c("subject", "age", "gender", "footedness", "handedness", "bilingual", "strongest.language")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(demo_dat$subject)[i] # find subject ID
  myrows <- which(demo_dat$subject==subname) # select rows for this subject
  tmp <- data.frame(demo_dat[myrows,])

@@

    demographics$subject[myrow] <- subname
    demographics$age[myrow] <- as.numeric(tmp[tmp$Question.Key == "Age",]$Response)
    demographics$gender[myrow] <- tmp[tmp$Question.Key == "Gender",]$Response
    demographics$footedness[myrow] <- tmp[tmp$Question.Key == "footedness",]$Response
    demographics$handedness[myrow] <- tmp[tmp$Question.Key == "categorical_hand",]$Response
    demographics$bilingual[myrow] <- tmp[tmp$Question.Key == "bilingual",]$Response
    demographics$strongest.language[myrow] <- tolower(tmp[tmp$Question.Key == "bilingual-text",]$Response)

    demographics$bilingual <- as.character(demographics$bilingual)
    demographics$bilingual <- gsub('. My strongest language is:', '', demographics$bilingual)
}

# Here we summarise the age broken down by handedness and gender. This will be important for any write up. 
#### NOTE. THIS IS NOT REPRESENTATIVE OF THE FINAL DATASET
# age
demographics %>%
    group_by(handedness, gender) %>%
    summarize(avg = mean(age), n = n(), sd = sd(age))
```

Now we import and append the Porta and Miles Test to the demographics data frame.

```{r sightednesss, warning=FALSE}
miles_dat_R <- read_csv("./Miles_R.csv", na = c("empty", "NA")) # read in data
miles_dat_L <- read_csv("./Miles_L.csv", na = c("empty", "NA")) # read in data
miles_dat <- rbind(miles_dat_R, miles_dat_L)
#porta_dat_R <- read_csv("./dat2/Porta_R.csv", na = c("empty", "NA")) # read in data
porta_dat_L <- read_csv("./Porta_L.csv", na = c("empty", "NA")) # read in data
porta_dat <- rbind(porta_dat_L)
# add test ID
miles_dat$test_ID <- "miles"
porta_dat$test_ID <- "porta"
# merge the two
sight_dat <- rbind(miles_dat, porta_dat)
# relabel 
sight_dat$subject <- sight_dat$`Participant Public ID`
sight_dat$subject <- as.factor(sight_dat$subject)

@@

# now merge sightedness and demographics
demographics <- merge(demographics, sightedness, by= "subject")
```

@@

```{r EHI, warning=FALSE}
EHI_R <- read_csv("./EHI_R.csv", na = "empty") # read in data
EHI_L <- read_csv("./EHI_L.csv", na = "empty") # read in data
EHI <- rbind(EHI_R, EHI_L)

@@

# score responses accoridng to the original scoring method in Oldifeld (1971)
EHI$right_hand <- 0
EHI$left_hand <- 0
for (r in 1:nrow(EHI)) {
  if(EHI$Response[r] == 1){
    EHI$right_hand[r]= 2 
  } else {
    if(EHI$Response[r] == 2){
      EHI$right_hand[r]= 1
    } else {
      if(EHI$Response[r] == 4){
        EHI$left_hand[r]= 1
      } else {
        if(EHI$Response[r] == 5){
          EHI$left_hand[r]= 2 
        } else {
          (EHI$right_hand[r]= 1) & (EHI$left_hand[r]= 1)
        }
      }
    }
  }
}

@@

# now create a data frame only for index 
EHI_index <- select(for_index2, "subject", "index_EHI")
# merge which demographics
demographics <- merge(demographics, EHI_index, by= "subject")

#now plot EHI in relation to handedness (modified by DB)
x<-ggplot(demographics, aes(x=index_EHI, y=subject, fill=subject)) +
  geom_dotplot(binaxis='y', stackdir='center') + xlim(-100, 100) +
  ggtitle("Handedness index on the Edinburgh Handedness Inventory")
x + theme_bw() + theme(legend.position = "none") + theme(axis.text.y = element_blank())

hist(demographics$index_EHI)
```

@@

```{r lexTALE, warning=FALSE}
# read data
lexTALE_L <- read_csv("./LexTale_L.csv", na = "empty")
lexTALE_R <- read_csv("./lexTale_R.csv", na = "empty")
lexTALE <- rbind(lexTALE_R, lexTALE_L)
# filter
lexTALE <- lexTALE[lexTALE$Attempt == 1,]
lexTALE <- lexTALE[lexTALE$display == "Task",]
# rename variables
lexTALE$subject <- lexTALE$`Participant Public ID`
lexTALE$accuracy <- lexTALE$Correct
lexTALE$condition <- lexTALE$ANSWER
# code as factors
lexTALE$subject <- as.factor(lexTALE$subject)
lexTALE$condition <- as.factor(lexTALE$condition)
  levels(lexTALE$condition) <- c("non-word", "word")

@@

# now create a data frame only for index 
lexTALE <- select(lexTale_score, "subject", "lexTALE")
# now merge with demographics
demographics <- merge(demographics, lexTALE, by= "subject")
# now write the full demographics 
write.csv(demographics, file = "./demographics_out.csv")

hist(demographics$lexTALE)
abline(v=80,col='red')

paste0("No. of participants below LexTALE cuttoff: ", sum(demographics$lexTALE < 80))

# remove those who didn't pass the LexTALE
demographics <- demographics[demographics$lexTALE >= 80,]
```

@@

Before doing anything with the data, we now pull in all the data from all nodes and combind them from a sub folder in the directory. Participants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

```{r readraw,include= FALSE, warning=FALSE}
# Set directory
for (d in 1:2){
  data_file <- "./day1_combined.csv"

  if (d ==2){
    data_file <-  "./day2_combined.csv"
  }
  #we will assign raw data to all_dat2 (day2), but then reassign to all_dat1 on first run through loop, so we have correct data for each day
all_dat2 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)
# cleaning - do operations that apply for all tasks
all_dat2 %>% filter(Attempt >= 1) # only answers 
# write as factors
all_dat2$subject <- as.factor(all_dat2$Participant.Public.ID)
all_dat2$RT <- as.numeric(as.character(all_dat2$Reaction.Time))
all_dat2$Trial <-as.numeric(all_dat2$Trial.Number) #useful if we want to restrict analysis to subset of trials so retain this variable
all_dat2$Task.Name <- as.factor(all_dat2$Task.Name)
all_dat2$word<-substring(all_dat2$word,66,(nchar(all_dat2$word)-4)) #remove format from written word stimuli
if (d==1)
{all_dat1 <- all_dat2}
}

# #remove cols from all_dat2 that contain 'branch' or 'randomiser'
# branchcols2 <- c(which(grepl('branch', names(all_dat2))=='TRUE'),which(grepl('randomiser', names(all_dat2))=='TRUE'))
# all_dat2 <- all_dat2[,-branchcols2]
# branchcols1 <- c(which(grepl('branch', names(all_dat1))=='TRUE'),which(grepl('randomiser', names(all_dat1))=='TRUE'))
# all_dat1 <- all_dat1[,-branchcols1]

#check that colnames are the same
# colcomp<-data.frame(colnames(all_dat2)[order(colnames(all_dat2))])
# colnames(colcomp)[1]<-'name1'
# ncol1 <-length(colnames(all_dat1))
# colcomp$name2<-NA
# colcomp$name2[1:ncol1]<-colnames(all_dat1)[order(colnames(all_dat1))]

#NB the colnames are identical but in different order, and all_dat1 has extra column of word_OG
#It is just conceivable the ordering might affect processing when columns are subsequently selected, so we'll reorder them to be consistent across sessions

@@

Another generic function that can be used for different tasks to make a laterality index and write it to the demographics file

```{r makeLIs, warning=FALSE}
# This function modifies a copy of the demographics file and returns it

@@

makeLI <- function(demogs,origdat,task,mycolname,polarity,mycutoff){
# we do this for accuracy first
# reformat data frame
keeps <- c("subject", "side", "X1") #X1 holds either accuracy or RT
LIdf <- origdat[keeps]
LIdf <- spread(LIdf, side, X1)
# calculate each participants' lat index
LIdf <- 
  LIdf %>% 
  group_by(subject) %>%
  mutate(thisLI= ((Right - Left)/(Right + Left))*100)

@@

# now create a data frame for lat index 
LIdf <- select(LIdf, "subject", "LI")
# now merge with demographics
demogs <- merge(demogs, LIdf, by= "subject")
lastcol<-length(colnames(demogs))
colnames(demogs)[lastcol]<-mycolname

return(demogs)
}
```

@@

```{r makeZ, warning=FALSE}
# This function modifies a copy of the demographics file and returns it

@@

makeZ <- function(demogs,origdat,task,mycolname,polarity,mycutoff){
# we do this for accuracy first
# reformat data frame
keeps <- c("subject", "side", "X1") #X1 holds either accuracy or RT
Zdf <- origdat[keeps]
Zdf <- spread(Zdf, side, X1)
# calculate each participants' lat index
Zdf <- 
  Zdf %>% 
  group_by(subject) %>%
  mutate(n= Left + Right, 
         pR= Right/n,
         pL= Left/n,
         Z= (pR-.5)/sqrt(pR*pL/n))
# remove infinite results
is.na(Zdf) <- do.call(cbind,lapply(Zdf, is.infinite))
Zdf <- na.omit(Zdf)

@@

# now create a data frame for lat index 
Zdf <- select(Zdf, "subject", "Z")
# now merge with demographics
demogs <- merge(demogs, Zdf, by= "subject")
lastcol<-length(colnames(demogs))
colnames(demogs)[lastcol]<-mycolname

return(demogs)
}
```

@@

CF_same <- 
  CF_same %>% 
  mutate(emotion_recog= ifelse(as.character(Response) == as.character(l_hemiface), 1, 0))
# calculate participants average
emotion_mean <- aggregate(FUN= mean, data= CF_same, emotion_recog~ subject)
# add this to demographics 
demographics <- merge(demographics, emotion_mean, by= "subject")
}
mycol <- length(colnames(demographics))
colnames(demographics)[(mycol-1):mycol]<-c('Emot.recog1','Emot.recog2')

plot(demographics$Emot.recog1,demographics$Emot.recog2)
abline(h=.75)
abline(v=.75)
text(.6,.8,'drop if <75% on either run',cex=.8)

#db modified to mark rather than drop cases with scores < .75
demographics$exclude <- 0 #default is to exclude

# mark the cases where there is less than 75% for identical stimuli
w <- union(which(demographics$Emot.recog1 < .75),which(demographics$Emot.recog2 < .75))
demographics$exclude[w]<-1 #code exclude = 1 if excluded because poor emot recog
```

@@

```{r domakeLI}
#We add Chimeric faces LIs to demographics for day 1 and then day2
task<-'Chimeric faces'
origdat<-chim_dat1
origdat$X1 <- origdat$accurate
mycolname<-"Day1_CF_acc_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

@@

#if we don't specify mycutoff, it will remain at level set for accuracy, which means it should not trap any RT values as they will all be much bigger
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

origdat<-chim_dat2
origdat$X1 <- origdat$accurate
mycolname<-"Day2_CF_acc_LI"
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"Day2_CF_RT_LI"
polarity <-1
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(demographics$Day1_CF_acc_LI,demographics$Day2_CF_acc_LI)
plot(demographics$Day1_CF_acc_LI,demographics$Day2_CF_acc_LI,main='Chimeric LI accuracy',col=(1+demographics$exclude))
abline(v=0)
abline(h=0)
text(-50,75,paste0('r = ',round(cor1,3)))
text(-50,65,'red will be excluded',cex=.85)
cor2<-cor(demographics$Day1_CF_RT_LI,demographics$Day2_CF_RT_LI)
plot(demographics$Day1_CF_RT_LI,demographics$Day2_CF_RT_LI,main='Chimeric LI RT',col=(1+demographics$exclude))
abline(v=0)
abline(h=0)
text(-10,20,paste0('r = ',round(cor2,3)))
text(-10,18,'red will be excluded',cex=.85)
```

```{r domakeZ}
#We add Chimeric faces LIs to demographics for day 1 and then day2
task<-'Chimeric faces'
origdat<-chim_dat1
origdat$X1 <- origdat$N
mycolname<-"Day1_CF_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeZ(demographics,origdat,task,mycolname,polarity,mycutoff)

origdat<-chim_dat2
origdat$X1 <- origdat$N
mycolname<-"Day2_CF_acc_Z"
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeZ(demographics,origdat,task,mycolname,polarity,mycutoff)
w<-which(demographics$excluded == 1)
demographics$Day1_CF_acc_Z[w] <- NA #removes zscores for excluded cases
demographics$Day2_CF_acc_Z[w] <- NA

RR<-intersect(which(demographics$Day1_CF_acc_Z>1.96),which(demographics$Day2_CF_acc_Z>1.96))
LL<-intersect(which(demographics$Day1_CF_acc_Z< -1.96),which(demographics$Day2_CF_acc_Z< -1.96))

cor1 <- cor(demographics$Day1_CF_acc_Z,demographics$Day2_CF_acc_Z,use='complete.obs',method='spearman')
plot(demographics$Day1_CF_acc_Z,demographics$Day2_CF_acc_Z,main='Chimeric Z accuracy')
abline(v=0)
abline(h=0)
abline(v = 1.96, col='red')
abline(v = -1.96, col='red')
abline(h = 1.96, col='red')
abline(h = -1.96, col='red')
text(-7,7.5,paste0('r = ',round(cor1,3)),cex=.8)
text(-2,-15,paste('Those in bottom quadrant\n bounded by red are consistently\n L-lateralised, N = ',length(LL)),cex=.8)
text(5,18,paste('Those in top quadrant\n bounded by red are consistently\n R-lateralised, N = ',length(RR)),cex=.8)
text(8,-8,'One outlier excluded',cex=.8)

@@

As we can see, we generally get a left side/right hemisphere advantage in our accuracy data. The RT advantage is not as expected. This may be due to the buttons being clicked on screen and the additional time penalty of moving the mouse.

@@

First, we judge participants' ability to correctly identify sounds using the "dichotic_same" dataframe. Participants are removed if their performance is below 75%.

@@

dichotic_same <- 
  dichotic_same %>% 
  mutate(sound_recog= ifelse(as.character(Response) == as.character(left_channel), 1, 0))
# calculate participants average
sound_mean <- aggregate(FUN= mean, data= dichotic_same, sound_recog~ subject)
# add this to demographics 
demographics <- merge(demographics, sound_mean, by= "subject")
}
mycol <- length(colnames(demographics))
colnames(demographics)[(mycol-1):mycol]<-c('sound.recog1','sound.recog2')

>>>>

---
title: "Online Laterality Battery Pre-Processing - short version"
author: "Adam Parker + Dorothy Bishop"
date: "updated April 22nd 2020"
output:
  pdf_document: default
  html_document: default
---
NB. 2 changes Apr 22nd by DB: 
a) all data now in a Data folder, so when reading need to specify that; 
b) *demographics* data frame renamed to *allsum*.

@@

This version updated on 1st Feb 2020 to make script more reproducible and compact by using functions for repeated operations.
Adapted in April 2020 for processing dataset on 300 people

The script pulls in csv files and processes them so that pre-registered analyses can be conducted (pre-registered on the OSF: ). First, questionnaires are analyses followed by chimeric, dichotic, visual half field and tapping tasks. This is done for time 1 and time 2. A "full" laterality index output is then appended to a file with the demographics. This is then used in subsequent R markdown scripts.

This script analyses the following: 
1) Questionnaires
  - demographics
  - Miles test
  - Porta test
  - Edinburgh Handedness Inventory
  - lexTALE
2) Chimeric face task
3) Dichotic listening
4) Rhyme detection task
5) Key Pressing Task (finger dexterity)
And then produces summary of demog and results for the final sample

@@

- demographics
- Miles test
- Porta test
- Edinburgh Handedness Inventory
- lexTALE

@@

#NB for original data setwd("~/Dropbox/Online LDT/300 Ps")
# libraries 
library(readr)
library(yarrr)
library(dplyr)
library(tidyr)
library(car)
library(ggplot2)
library(knitr)
require(readr)  # for read_csv()
require(dplyr)  # for mutate()
require(tidyr)  # for unnest()
require(purrr)  # for map(), reduce()
library(GGally)
require(stringr) # string matching
library(ggpirate)
library(corrr) #easier correlations!
```

@@

```{r demograhics, warning=FALSE}
demo_dat_R <- read_csv("./data/demographics_R.csv", na = c("empty", "NA")) # read in data
demo_dat_L <- read_csv("./data/demographics_L.csv", na = c("empty", "NA")) # read in data
demo_dat <- rbind(demo_dat_R, demo_dat_L)
# relabel 
demo_dat$subject <- demo_dat$`Participant Public ID`
demo_dat$subject <- as.factor(demo_dat$subject)
# count and print subjects
nsub <-length(levels(demo_dat$subject))
print(paste0("subjects: ", nsub))

# create data frame
allsum <- data.frame(matrix(ncol = 7, nrow = nsub))
  colnames(allsum) <- c("subject", "age", "gender", "footedness", "handedness", "bilingual", "strongest.language")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(demo_dat$subject)[i] # find subject ID
  myrows <- which(demo_dat$subject==subname) # select rows for this subject
  tmp <- data.frame(demo_dat[myrows,])

@@

    allsum$subject[myrow] <- subname
    allsum$age[myrow] <- as.numeric(tmp[tmp$Question.Key == "Age",]$Response)
    allsum$gender[myrow] <- tmp[tmp$Question.Key == "Gender",]$Response
    allsum$footedness[myrow] <- tmp[tmp$Question.Key == "footedness",]$Response
    allsum$handedness[myrow] <- tmp[tmp$Question.Key == "categorical_hand",]$Response
    allsum$bilingual[myrow] <- tmp[tmp$Question.Key == "bilingual",]$Response
    allsum$strongest.language[myrow] <- tolower(tmp[tmp$Question.Key == "bilingual-text",]$Response)

    allsum$bilingual <- as.character(allsum$bilingual)
    allsum$bilingual <- gsub('. My strongest language is:', '', allsum$bilingual)
}

# Here we summarise the age broken down by handedness and gender. This will be important for any write up. 
#### NOTE. THIS IS NOT REPRESENTATIVE OF THE FINAL DATASET
# age
allsum %>%
    group_by(handedness, gender) %>%
    summarize(avg = mean(age), n = n(), sd = sd(age))

#create a numerical code for handedness - can be used in plots to distinguish handedness groups (eg if pch set to handpch, get circles and triangles)
allsum$handpch <-1
allsum$handpch[allsum$handedness=='Left'] <-2
```

Now we import and append the Porta and Miles Test to the allsum data frame.

```{r sightednesss, warning=FALSE}
miles_dat_R <- read_csv("./data/Miles_R.csv", na = c("empty", "NA")) # read in data
miles_dat_L <- read_csv("./data/Miles_L.csv", na = c("empty", "NA")) # read in data
miles_dat <- rbind(miles_dat_R, miles_dat_L)
#porta_dat_R <- read_csv("./dat2/Porta_R.csv", na = c("empty", "NA")) # read in data
porta_dat_L <- read_csv("./data/Porta_L.csv", na = c("empty", "NA")) # read in data
porta_dat <- rbind(porta_dat_L)
# add test ID
miles_dat$test_ID <- "miles"
porta_dat$test_ID <- "porta"
# merge the two
sight_dat <- rbind(miles_dat, porta_dat)
# relabel 
sight_dat$subject <- sight_dat$`Participant Public ID`
sight_dat$subject <- as.factor(sight_dat$subject)

@@

# now merge sightedness and allsum
allsum <- merge(allsum, sightedness, by= "subject")
```

@@

```{r EHI, warning=FALSE}
EHI_R <- read_csv("./data/EHI_R.csv", na = "empty") # read in data
EHI_L <- read_csv("./data/EHI_L.csv", na = "empty") # read in data
EHI <- rbind(EHI_R, EHI_L)

@@

# score responses according to the original scoring method in Oldifeld (1971)
EHI$right_hand <- 0
EHI$left_hand <- 0
for (r in 1:nrow(EHI)) {
  if(EHI$Response[r] == 1){
    EHI$right_hand[r]= 2 
  } else {
    if(EHI$Response[r] == 2){
      EHI$right_hand[r]= 1
    } else {
      if(EHI$Response[r] == 4){
        EHI$left_hand[r]= 1
      } else {
        if(EHI$Response[r] == 5){
          EHI$left_hand[r]= 2 
        } else {
          (EHI$right_hand[r]= 1) & (EHI$left_hand[r]= 1)
        }
      }
    }
  }
}

@@

# now create a data frame only for index 
EHI_index <- select(for_index2, "subject", "index_EHI")
# merge with allsum
allsum <- merge(allsum, EHI_index, by= "subject")

#now plot EHI in relation to handedness (modified by DB)

eplot<-ggplot(allsum, aes(handpch,index_EHI, colour=gender)) +
  geom_jitter(width=.25,height=2)+
  ggtitle("Handedness index on the Edinburgh Handedness Inventory")
eplot
#This is not perfect, but gives and idea to check EHI is valid etc.

ggplot(allsum, aes(x=index_EHI)) + geom_density(aes(group=handpch,fill=handpch))

#NB one supposed right-hander has index_EHI around -50. 
#I will recode this one, as this seems an error.

oddcase <- intersect(which(allsum$index_EHI<0),which(allsum$handedness=='Right'))
allsum$handedness[oddcase] <- 'Left'
allsum$handpch[oddcase] <- 1
```

@@

```{r lexTALE, warning=FALSE}
# read data
lexTALE_L <- read_csv("./data/LexTale_L.csv", na = "empty")
lexTALE_R <- read_csv("./data/lexTale_R.csv", na = "empty")
lexTALE <- rbind(lexTALE_R, lexTALE_L)
# filter
lexTALE <- lexTALE[lexTALE$Attempt == 1,]
lexTALE <- lexTALE[lexTALE$display == "Task",]
# rename variables
lexTALE$subject <- lexTALE$`Participant Public ID`
lexTALE$accuracy <- lexTALE$Correct
lexTALE$condition <- lexTALE$ANSWER
# code as factors
lexTALE$subject <- as.factor(lexTALE$subject)
lexTALE$condition <- as.factor(lexTALE$condition)
  levels(lexTALE$condition) <- c("non-word", "word")

@@

# now create a data frame only for index 
lexTALE <- select(lexTale_score, "subject", "lexTALE")
# now merge with allsum
allsum <- merge(allsum, lexTALE, by= "subject")
# now write the full allsum 
write.csv(allsum, file = "./allsum_out.csv")

hist(allsum$lexTALE)
abline(v=80,col='red')

paste0("No. of participants below LexTALE cuttoff: ", sum(allsum$lexTALE < 80))

# remove those who didn't pass the LexTALE
allsum <- allsum[allsum$lexTALE >= 80,]
```

@@

Before doing anything with the data, we now pull in all the data from all nodes and combine them from a sub folder in the directory. Participants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

```{r readraw,include= FALSE, warning=FALSE}
# Set directory
for (d in 1:2){
  data_file <- "./data/day1_combined.csv"

  if (d ==2){
    data_file <-  "./data/day2_combined.csv"
  }
  #we will assign raw data to all_dat2 (day2), but then reassign to all_dat1 on first run through loop, so we have correct data for each day
all_dat2 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)
# cleaning - do operations that apply for all tasks
all_dat2 %>% filter(Attempt >= 1) # only answers 
# write as factors
all_dat2$subject <- as.factor(all_dat2$Participant.Public.ID)
all_dat2$RT <- as.numeric(as.character(all_dat2$Reaction.Time))
all_dat2$Trial <-as.numeric(all_dat2$Trial.Number) #useful if we want to restrict analysis to subset of trials so retain this variable
all_dat2$Task.Name <- as.factor(all_dat2$Task.Name)
all_dat2$word<-substring(all_dat2$word,66,(nchar(all_dat2$word)-4)) #remove format from written word stimuli
if (d==1)
{all_dat1 <- all_dat2}
}

@@

Another generic function that can be used for different tasks to make a laterality index and write it to the allsum file

```{r makeLIs, warning=FALSE}
# This function modifies a copy of the allsum file and returns it

@@

makeLI <- function(allsums,origdat,task,mycolname,polarity,mycutoff){
# we do this for accuracy first
# reformat data frame
keeps <- c("subject", "side", "X1") #X1 holds either accuracy or RT
LIdf <- origdat[keeps]
LIdf <- spread(LIdf, side, X1)
# calculate each participants' lat index
LIdf <- 
  LIdf %>% 
  group_by(subject) %>%
  mutate(thisLI= ((Right - Left)/(Right + Left))*100)

@@

# now create a data frame for lat index 
LIdf <- select(LIdf, "subject", "LI")
# now merge with allsum
allsums <- merge(allsums, LIdf, by= "subject")
lastcol<-length(colnames(allsums))
colnames(allsums)[lastcol]<-mycolname

return(allsums)
}
```

@@

```{r makeZ, warning=FALSE}
# This function modifies a copy of the allsum file and returns it

@@

makeZ <- function(allsums,origdat,task,mycolname,polarity,mycutoff){
# we do this for accuracy first
# reformat data frame
keeps <- c("subject", "side", "X1") #X1 holds either accuracy or RT
Zdf <- origdat[keeps]
Zdf <- spread(Zdf, side, X1)
# calculate each participants' lat index
Zdf <- 
  Zdf %>% 
  group_by(subject) %>%
  mutate(n= Left + Right, 
         pR= Right/n,
         pL= Left/n,
         Z= (pR-.5)/sqrt(pR*pL/n))
# remove infinite results
is.na(Zdf) <- do.call(cbind,lapply(Zdf, is.infinite))
Zdf <- na.omit(Zdf)

@@

# now create a data frame for lat index 
Zdf <- select(Zdf, "subject", "Z")
# now merge with allsum
allsums <- merge(allsums, Zdf, by= "subject")
lastcol<-length(colnames(allsums))
colnames(allsums)[lastcol]<-mycolname

return(allsums)
}
```

@@

CF_same <- 
  CF_same %>% 
  mutate(emotion_recog= ifelse(as.character(Response) == as.character(l_hemiface), 1, 0))
# calculate participants average
emotion_mean <- aggregate(FUN= mean, data= CF_same, emotion_recog~ subject)
# add this to allsum 
allsum <- merge(allsum, emotion_mean, by= "subject")
}
mycol <- length(colnames(allsum))
colnames(allsum)[(mycol-1):mycol]<-c('Emot.recog1','Emot.recog2')

plot(jitter(allsum$Emot.recog1,10),jitter(allsum$Emot.recog2,10),pch=3,cex=.6)
abline(h=.75)
abline(v=.75)
text(.6,.8,'drop if <75% on either run',cex=.8)

#db modified to mark rather than drop cases with scores < .75
allsum$exclude <- 0 #default is to exclude

# mark the cases where there is less than 75% for identical stimuli
w <- union(which(allsum$Emot.recog1 < .75),which(allsum$Emot.recog2 < .75))
allsum$exclude[w]<-1 #code exclude = 1 if excluded because poor emot recog
```

@@

```{r domakeLI}
#We add Chimeric faces LIs to allsum for day 1 and then day2
task<-'Chimeric faces'
origdat<-chim_dat1
origdat$X1 <- origdat$accurate
mycolname<-"Day1_CF_acc_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

@@

#if we don't specify mycutoff, it will remain at level set for accuracy, which means it should not trap any RT values as they will all be much bigger
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-chim_dat2
origdat$X1 <- origdat$accurate
mycolname<-"Day2_CF_acc_LI"
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"Day2_CF_RT_LI"
polarity <-1
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(allsum$Day1_CF_acc_LI,allsum$Day2_CF_acc_LI)
plot(allsum$Day1_CF_acc_LI,allsum$Day2_CF_acc_LI,main='Chimeric LI accuracy',col=(1+allsum$exclude))
abline(v=0)
abline(h=0)
text(-50,75,paste0('r = ',round(cor1,3)))
text(-50,65,'red will be excluded',cex=.85)
cor2<-cor(allsum$Day1_CF_RT_LI,allsum$Day2_CF_RT_LI)
plot(allsum$Day1_CF_RT_LI,allsum$Day2_CF_RT_LI,main='Chimeric LI RT',col=(1+allsum$exclude))
abline(v=0)
abline(h=0)
text(-10,20,paste0('r = ',round(cor2,3)))
text(-10,18,'red will be excluded',cex=.85)
```

Here we use the Z score version. This computes for each subject whether they have a different proportion correct for L and R

```{r domakeZ}
#We add Chimeric faces LIs to allsum for day 1 and then day2
task<-'Chimeric faces'
origdat<-chim_dat1
origdat$X1 <- origdat$N
mycolname<-"Day1_CF_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-chim_dat2
origdat$X1 <- origdat$N
mycolname<-"Day2_CF_acc_Z"
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
w<-which(allsum$excluded == 1)
allsum$Day1_CF_acc_Z[w] <- NA #removes zscores for excluded cases
allsum$Day2_CF_acc_Z[w] <- NA

RR<-intersect(which(allsum$Day1_CF_acc_Z>1.96),which(allsum$Day2_CF_acc_Z>1.96))
LL<-intersect(which(allsum$Day1_CF_acc_Z< -1.96),which(allsum$Day2_CF_acc_Z< -1.96))

cor1 <- cor(allsum$Day1_CF_acc_Z,allsum$Day2_CF_acc_Z,use='complete.obs',method='spearman')
plot(allsum$Day1_CF_acc_Z,allsum$Day2_CF_acc_Z,main='Chimeric Z accuracy')
abline(v=0)
abline(h=0)
abline(v = 1.96, col='red')
abline(v = -1.96, col='red')
abline(h = 1.96, col='red')
abline(h = -1.96, col='red')
text(-7,7.5,paste0('Spearman r = ',round(cor1,3)),cex=.8)
text(-2,-15,paste('Those in bottom quadrant\n bounded by red are consistently\n L-lateralised, N = ',length(LL)),cex=.8)
text(5,18,paste('Those in top quadrant\n bounded by red are consistently\n R-lateralised, N = ',length(RR)),cex=.8)
#text(8,-8,'One outlier excluded',cex=.8)
#Removal of the outlier does not have much influence on rs

@@

As we can see, we generally get a left side/right hemisphere advantage in our accuracy data. We would not expect a RT advantage for this type of measure, where the response is to a single stimulus with conflicting information from the 2 sides.

@@

First, we judge participants' ability to correctly identify sounds using the "dichotic_same" dataframe. Participants are marked as excluded if their performance is below 75%.

@@

dichotic_same <- 
  dichotic_same %>% 
  mutate(sound_recog= ifelse(as.character(Response) == as.character(left_channel), 1, 0))
# calculate participants average
sound_mean <- aggregate(FUN= mean, data= dichotic_same, sound_recog~ subject)
# add this to allsum 
allsum <- merge(allsum, sound_mean, by= "subject")
}
mycol <- length(colnames(allsum))
colnames(allsum)[(mycol-1):mycol]<-c('sound.recog1','sound.recog2')

# once again, mark rather than remove those with poor performance on identical stimuli
w<-union(which(allsum$sound.recog1 < .75), which(allsum$sound.recog2 < .75))
ww<-intersect(which(allsum$exclude==1),w)
#ww contains people who were excluded on both dichotic and faces. 
# code exclude = 2 for the cases where there less than 75% for identical stimuli
allsum$exclude[w]<-2 #code exclude = 2 if excluded because poor dichotic
allsum$exclude[ww]<-12 #code for those excluded on both chimeric and dichotic
t <- table(allsum$exclude) 
rownames(t)<-c('include','exclude CF','exclude dich','exclude CF+dich')
t
```
